\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-04-11 paper/main.tex
Results changes and Discussion intro

Proteins, neurotransmitters and other molecular substances
intrinsically diffuse within the brain ventricles, but are also
advected by the flow of CSF. A natural question is which of these
mechanisms that dominate for any given type of solute. To address this
question, we simulate the distribution and evolution of three
different diffusing solutes injected within a region of interest (ROI 1) in the
dorsal region of the middle ventricle (\Crefpanels{fig:fig2}{a}),
subject to the CSF flow induced by motile cilia and cardiac pressure
pulsations. In particular, we consider three different molecular
diffusion coefficients, corresponding to a large extracellular vesicle (EV, $D_1 = 2.17
\times 10^{-12}$ m$^2$/s), a 90 kDa Starmaker-Green Fluorescent Protein
(STM-GFP, $D_2 = 5.75 \times 10^{-11}$ m$^2$/s),
and a 26 kDa photoconvertible protein
Dendra2 ($D_3 = 1.15 \times 10^{-10}$ m$^2$/s).

@@

Ciliary motion is integral to physiological CSF flow,
and motile-cilia deficiency has been
associated with defective brain development and various
pathologies in zebrafish, rodents and humans~\cite{Eichele2020Cilia-drivenVentricle,
Olstad2019CiliaryDevelopment, Faubel2016Cilia-basedVentricles,
Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
Sawamoto2006NewBrain, Yoshiba2014RolesSymmetry,
Hirokawa2006NodalAsymmetry, Youn2018PrimaryDiseases,
Ma2024CiliaDisease, Ibanez-Tallon2002LossHydrocephalus}.  As observed both experimentally and
computationally, absence of motile cilia alters CSF flow patterns.  We
next investigate the implications of these changes on brain ventricular
transport. We experimentally measured the movement of proteins by
monitoring photoconverted Dendra2 signal in control zebrafish and
motile-cilia mutant \emph{schmalhans} (\emph{smh}) zebrafish.  In ROIs
2 and 4, which quantify Dendra2 distribution downstream of the
photoconversion site, we observed a slower increase of fluorescence
intensity (\Crefpanels{fig:fig4}{a}): the mean times-to-threshold in
the \emph{smh} mutants were $45\%$ (ROI 2) and $25\%$ (ROI 4) higher
compared to controls (\Crefpanels{fig:fig4}{b}). We found no
significant differences between \emph{smh} mutants and controls in the
Dendra2 distribution in ROIs 3, 5, and 6.

In the computational model, we represent the \emph{smh} mutants by
considering transport of Dendra2 with CSF flow driven by cardiac
pressure pulsations only, representing the absence of motile
cilia. Comparing the in silico mutants with transport predictions in
the presence of both cardiac pulsations and motile cilia, we observe
noticeably delayed dynamics in ROIs 2, 4 and 5, but little change for
ROIs 3 and 6 (\Crefpanels{fig:fig4}{c}). Compared to experimental
data, the control and mutant scenarios differed more for the
simulations, with the times-to-threshold in ROI 2 and 4 being 81\% and
28\% higher in the absence of cilia-mediated flow. Moreover, the
predicted mean concentrations in ROI 5 deviate between the two
scenarios, which is not the case in the experiments. Altogether, our
findings identify that, for smaller proteins such as Dendra2,
diffusion plays an important role for solute movement to regions
upstream in the cilia-mediated flow fields or to distant ventricles,
while the cilia affect the downstream mean concentrations.
\begin{figure}[H]
    \centering
    \includegraphics[width=0.96\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $\Delta F(t)$
    in each region of interest (ROI) for control (black lines)
    and \emph{smh} mutant (red lines) zebrafish. 
    Bold lines are mean values of the cohorts and shaded regions cover one standard deviation.
    \textbf{b.} Times when $\Delta F(t)$ first exceeds a threshold value,
    equal to 0.25 (ROIs 2--4) or 0.10 (ROIs 5, 6), for controls (gray dots) and
    \emph{smh} mutants (red dots).
    The thick black lines denote mean values. The $p$-values
    were calculated using a Mann-Whitney U test with 95\% confidence level,
    under the null hypothesis that the two distributions of control and mutant data are 
    from the same population.
    \textbf{c.} Mean concentration $\cbar(t)$ in each ROI,
    simulated using the baseline model (black lines) and the cardiac-induced/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time
    $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in panels \textbf{a} and \textbf{c} the vertical axes have
    different limits in each plot.}
    \label{fig:fig4}
\end{figure}

@@

\section{Discussion}
In this study, we have presented a computational model of CSF flow and
solute transport within brain ventricles.  We used the larval
zebrafish brain ventricles since their geometry and flow mechanisms are
well described, thereby allowing for validation of the simulation
results against experimental measurements. Our computational framework
can be adapted to other ventricular systems, once the flow parameters
of such are adequately resolved. Here, we assumed that
intraventricular CSF flow was governed by the Stokes equations. By
imposing tangential traction and normal pressure acting on the CSF, we
were able to reproduce flow features observed experimentally. In
particular, vortex-structured flow mediated by the cilia and pulsatile
flow mediated by the cardiac cycle emerged as key flow characteristics.

>>>>

Proteins, neurotransmitters and other molecules
intrinsically diffuse within the brain ventricles, but are also
advected by the flow of CSF. A natural question is which of these
mechanisms dominates for any given type of solute. To address this
question, we simulate the distribution and evolution of three
different solutes injected within a region of interest (ROI 1) in the
dorsal region of the middle ventricle (\Crefpanels{fig:fig2}{a}) and
subject to the CSF flow induced by motile cilia and cardiac pressure
pulsations. In particular, we consider three different molecular
diffusion coefficients, corresponding to a large extracellular vesicle (EV, $D_1 = 2.17
\times 10^{-12}$ m$^2$/s), a 90 kDa Starmaker-Green Fluorescent Protein
(STM-GFP, $D_2 = 5.75 \times 10^{-11}$ m$^2$/s),
and a 26 kDa photoconvertible protein
Dendra2 ($D_3 = 1.15 \times 10^{-10}$ m$^2$/s).

@@

Ciliary motion is integral to physiological CSF flow,
and motile-cilia deficiency has been
associated with defective brain development and various
pathologies in zebrafish, rodents and humans~\cite{Eichele2020Cilia-drivenVentricle,
Olstad2019CiliaryDevelopment, Faubel2016Cilia-basedVentricles,
Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
Sawamoto2006NewBrain, Yoshiba2014RolesSymmetry,
Hirokawa2006NodalAsymmetry, Youn2018PrimaryDiseases,
Ma2024CiliaDisease, Ibanez-Tallon2002LossHydrocephalus}.  As observed both experimentally and
computationally, absence of motile cilia alters CSF flow patterns
\Crefpanels{fig:fig1}{g}~\cite{Olstad2019CiliaryDevelopment}. We
next investigate the implications of these changes on brain ventricular
transport. We experimentally measured the movement of proteins by
monitoring photoconverted Dendra2 signal in control zebrafish and
motile-cilia mutant \emph{schmalhans} (\emph{smh}) zebrafish.  In ROIs
2 and 4, which quantify Dendra2 distribution downstream of the
photoconversion site, we observed a slower increase of fluorescence
intensity (\Crefpanels{fig:fig4}{a}): the mean times-to-threshold in
the \emph{smh} mutants were $45\%$ (ROI 2) and $25\%$ (ROI 4) higher
compared to controls (\Crefpanels{fig:fig4}{b}). We found no
significant differences between \emph{smh} mutants and controls in the
Dendra2 distribution in ROIs 3, 5, and 6.

In the computational model, we represent the \emph{smh} mutants by
considering transport of Dendra2 with CSF flow driven by cardiac
pressure pulsations only, representing the absence of motile
cilia. Comparing the in silico mutants with transport predictions in
the presence of both cardiac pulsations and motile cilia, we observe
noticeably delayed dynamics in ROIs 2, 4 and 5, but little change for
ROIs 3 and 6 (\Crefpanels{fig:fig4}{c}). Compared to experimental
data, the control and mutant scenarios differed more for the
simulations, with the times-to-threshold in ROI 2 and 4 being 81\% and
28\% higher in the absence of cilia-mediated flow. Moreover, the
predicted mean concentrations in ROI 5 deviate between the two
scenarios, while in the experiments the difference in
fluorescence intensities is not significant.
Altogether, our findings identify that,
for smaller proteins such as Dendra2,
diffusion plays an important role for solute movement to regions
upstream in the cilia-mediated flow fields or to distant ventricles,
while the cilia affect the downstream mean concentrations.
\begin{figure}[H]
    \centering
    \includegraphics[width=0.96\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $\Delta F(t)$
    in each region of interest (ROI) for control (black lines)
    and \emph{smh} mutant (red lines) zebrafish. 
    Bold lines are mean values of the cohorts and shaded regions cover one standard deviation.
    \textbf{b.} Times when $\Delta F(t)$ first exceeds a threshold value,
    equal to 0.25 (ROIs 2--4) or 0.10 (ROIs 5, 6), for controls (gray dots) and
    \emph{smh} mutants (red dots).
    The thick black lines denote mean values. The $p$-values
    were calculated using a Mann-Whitney U test with 95\% confidence level,
    under the null hypothesis that the two distributions of control and mutant data are 
    from the same population.
    \textbf{c.} Mean concentration $\cbar(t)$ in each ROI,
    simulated using the baseline model (black lines) and the cardiac-induced/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time
    $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in panels \textbf{a} and \textbf{c} the vertical axes have
    different limits in each plot.}
    \label{fig:fig4}
\end{figure}

@@

\section{Discussion}
In this study, we have presented a computational model of CSF flow and
solute transport within brain ventricles.  We used the larval
zebrafish brain ventricles since their geometry and flow mechanisms are
well described, thereby allowing for validation of the simulation
results against experimental measurements. Our computational framework
can be adapted to other ventricular systems, once their flow parameters
are adequately resolved. Here, we assumed that
intraventricular CSF flow was governed by the Stokes equations. By
imposing tangential traction and normal pressure acting on the CSF, we
were able to reproduce flow features observed experimentally. In
particular, vortex-structured flow mediated by the cilia and pulsatile
flow mediated by the cardiac cycle emerged as key flow characteristics.
